### 2023-08-14 appendix.tex
up

We now consider the case where $\Safe = \uparrow \Safe$ and $\Bad$ is not the complement of 
$\Safe$ and we'd want $\Bad$ to have the same closure properties as $\Safe$.
We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the subcase 
$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction, since two upward-closed sets have not an empty intersection. We thus consider instead the case where $\Bad$ is as upward-closed
as it can be without intersecting with $\Safe$,
i.e. 
such that there exists an upward-closed set $U$
such that $\Bad = U \setminus \Safe$.

@@

Home-state is undecidable for Minsky machine with $3+$ counters.
This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
Thus {\sc Home-state} is undecidable for Minsky machines. Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}. Hence {\sc Resilience} is undecidable for Minsky machines.

Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
% Let us recall more formally what a reset-VASS is.
Reset-VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
\begin{definition}
A {\em reset-VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

@@

It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.

@@

We are now interested in the 
%following
 undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset VASS, which consists in, given a
Reset VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
which we then reduce to {\sc State-resilience}.
A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
live. This leads to the following problem.

\problemx{$t$-liveness}
{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
{Is $t$ live in $s_0$ ? \\}

@@

\begin{proposition}\label{reductions}
{\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
\end{proposition}

\begin{proof}
We reformulate $t$-liveness 
in a 
% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
Reset $d$-VASS $(Q,T)$ 
with initial configuration $s_0$
as the following formula
\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
where
$U_t=\uparrow pre(t)$ 
is the upward closure of the preconditions to fire transition $t$.  
The problem reduces itself to {\sc State-resilience}
where $\Safe = U_t$.
\end{proof}

@@

\mathieu{À refaire avec des Reset VASS}

@@

\begin{proof}
If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.

@@

Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
% the following:
%
%
{\sc %$ Reset $d$-VASS 
$t$-liveness} is undecidable.

Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility. Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility. %This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

>>>>

We now consider the case where $\Safe = \uparrow \Safe$ and $\Bad$ is not the complement of 
$\Safe$ and we'd want $\Bad$ to have the same closure properties as $\Safe$.
We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the conditions 
$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ would lead to a contradiction, since two upward-closed sets have not an empty intersection. We consider instead the case where $\Bad$ is as upward-closed
as it can be without intersecting with $\Safe$,
i.e. 
such that there exists an upward-closed set $U$
such that $\Bad = U \setminus \Safe$.

@@

Home-state is undecidable for Minsky machine with $3+$ counters.
This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
Thus {\sc Home-state} is undecidable for Minsky machines. Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}. Hence {\sc Resilience} is undecidable for Minsky machines.

Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
% Let us recall more formally what a reset-VASS is.
Reset-VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
\begin{definition}
A {\em reset-VASS} in dimension $d$ %(reset-VASS for short)
 is a finite 
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

@@

It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.

@@

We are now interested in the 
%following
 undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
which we then reduce to {\sc State-resilience}.
A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
live. This leads to the following problem.

\problemx{$t$-liveness}
{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
{Is $t$ live in $s_0$ ? \\}

@@

\begin{proposition}\label{reductions}
{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
\end{proposition}

\begin{proof}
We reformulate $t$-liveness 
in a 
% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
reset-VASS $(Q,T)$ 
with initial configuration $s_0$
as the following formula
\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
where
$U_t=\uparrow pre(t)$ 
is the upward closure of the preconditions to fire transition $t$.  
The problem reduces itself to {\sc State-resilience}
where $\Safe = U_t$.
\end{proof}

@@

@@

\begin{proof}

We wish to determine if $q(\textbf{0})$ is reachable in a reset-VASS $V_1$
with
initial vector $\textbf{u}$ and initial state $q$.
To do so, we construct a reset-VASS $V_2$
with initial vector $\textbf{v}$ and initial state $q'$
which is live if and only if 
$\textbf{0}$
is reachable from $\textbf{u}$.

The new reset-VASS is constructed by adding two new places
and
$|V_1|+2$ new transitions.

We first modify all transitions to include $r_1$ both as an input and an output

\mathieu{Ok non pas comme ça}

@@

Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
% the following:
%
%
{\sc %$ reset-VASS 
$t$-liveness} is undecidable.

Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility. Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility. %This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
